---
title: "Project1"
author: "Zhijun Liu"
date: "`r Sys.Date()`"
output:  
  rmarkdown::github_document:  
    toc: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,message = FALSE,warning = FALSE)
```

```{r,echo=FALSE}
# prepare for the packages
library(httr)
library(jsonlite)
library(tidyverse)
library(DT)
library(knitr)
```

# Abstract
*What did I do + brief results*

# Introduction
## Description of JSON Data
### What it is?

JSON (JavaScript Object Notation) is an open standard file format, and data interchange format, that uses human-readable text to store and transmit data objects consisting of attribute–value pairs and array data types (or any other serializable value). JSON is a language-independent data format. It was derived from JavaScript, but many modern programming languages include code to generate and parse JSON-format data.(https://en.wikipedia.org/wiki/JSON)

### Where does it get used?

JSON is commonly used for transmitting data in web applications (e.g., sending some data from the server to the client, so it can be displayed on a web page, or vice versa). And It’s also becoming an increasingly common format for database migration from modern apps (such as MixPanel, SalesForce, and Shopify) over to SQL databases.(https://developer.mozilla.org/en-US/docs/Learn/JavaScript/Objects/JSON)

### Why is it a good way to store data?

JSON is perfect for storing temporary data that’s consumed by the entity that creates the data. A good example is user-generated data such as filling out a form or information exchange between an API and an app. Stored JSON data must be text but this means JSON can be used as a data format for any programming language, providing a high level of interoperability. It also means data stored in JSON files are easily sent between servers. This is a bonus for database migration.(https://blog.sqlizer.io/posts/json-store-data/)

## Discussion on R packages for reading JSON data

- `rjson`: This does not use S4/S3 methods and so is not readily extensible, but still useful. Unfortunately, it does not used vectorized operations and so is too slow for non-trivial data. Similarly, for reading JSON data into R, it is somewhat slow and so does not scale to large data, should this be an issue.

- `RJSONIO`: This is an alternative to `rjson` package. Originally, that was too slow for converting large R objects to JSON and was not extensible. rjson's performance is now similar to this package, and perhaps slightly faster in some cases. The two packages intentionally share the same basic interface. This package (`RJSONIO`) has many additional options to allow customizing the generation and processing of JSON content. This package uses libjson rather than implementing yet another JSON parser. 

- `jsonlite`: This is a fork of the `RJSONIO` package. It builds on the parser from `RJSONIO` but implements a different mapping between R objects and JSON strings. The C code in this package is mostly from the `RJSONIO` Package, the R code has been rewritten from scratch. In addition to drop-in replacements for `fromJSON` and `toJSON`, the package has functions to serialize objects. Furthermore, the package contains a lot of unit tests to make sure that all edge cases are encoded and decoded consistently for use with dynamic data in systems and applications. Thus, I choose `jsonlite`.

# NHL & NHL Data
## What is NHL?

The National Hockey League (NHL) is a professional ice hockey league in North America, currently comprising 31 teams: 24 in the United States and seven in Canada. The NHL is considered to be the premier professional ice hockey league in the world, and one of the major professional sports leagues in the United States and Canada. The Stanley Cup, the oldest professional sports trophy in North America, is awarded annually to the league playoff champion at the end of each season.(https://en.wikipedia.org/wiki/National_Hockey_League)

The NHL divides the 31 teams into two conferences: the Eastern Conference and the Western Conference. Each conference is split into two divisions: the Eastern Conference contains 16 teams (eight per division), while the Western Conference has 15 teams (seven in the Central Division and eight in the Pacific Division). The current alignment has existed since the 2017–18 season.(https://en.wikipedia.org/wiki/Category:National_Hockey_League_teams)

## View the NHL data

```{r}
# write a function to get specific url
url <- function(kind,ID=NULL){
  base_url <- "https://records.nhl.com/site/api/franchise"
  if(kind == "base"){
    return("https://records.nhl.com/site/api/franchise")}
  else if(kind == "TeamTotal"){
    return( paste(base_url, "team","totals", sep="-"))
  }else{
    return(paste0(paste(base_url,kind,"records",sep="-"),"?",paste("cayenneExp","franchiseId",ID,sep="=")))
  }
}
# write a function to get data from specific url
GetData <- function(url){
  url %>%
    GET(.) %>%
    # turn it into JSON text form
    content(as="text") %>%
    # convert it to a list
    fromJSON(flatten=TRUE) 
}
```

```{r}
# call above two functions together to get specific data
franchiseRaw <- GetData(url("base"))$data
teamtotalRaw <- GetData(url("TeamTotal"))$data
# write a function to combine all the id(1-38) of specific table
CData <- function(kind){
  y <- NULL
  for(i in 1:38){
    x <- GetData(url(kind,ID=as.character(i)))$data
    y <- rbind(x,y)
  }
  return(y)
}
# get the season table of all id
seasonRaw <- CData("season")
# get the goalie table of all id
goalieRaw <- CData("goalie")
# get the skater table of all id
skaterRaw <- CData("skater")
```

```{r}
# franchiesData is for looking up franchise id
franchiseData <-franchiseRaw %>% 
  select(id,contains("Name")) %>% 
  # the id in this table means franchiseId
  rename("franchiseId"=id)

# teamtotalData is for getting the WinPct to generate the futureWin
teamtotalData <- teamtotalRaw %>% 
  select(franchiseId ,wins,gamesPlayed,activeFranchise) %>%
  group_by(franchiseId) %>%
  # sum the number of total wins and the number of total games
  summarise(sumWins=sum(wins),sumGames=sum(gamesPlayed),activeFranchise = mean(activeFranchise)) %>%
  # get the WinPct variable by divide above two variables, and round into 2 decimal
  mutate(winPct=round((sumWins/sumGames),2))

# seasonData is for getting the WinPct mfwinRation to generate the steadyWin
seasonData <- seasonRaw %>% 
  select(14,12,mostWins) %>%
  # create a new variable called mfwinRation, calculated by the ration between mostWins and fewestWins
  mutate(mfWinRatio = round(mostWins/fewestWins,2))

# goalieData is for knowing the difference in goalie of specific franchise considering teamtotalData and seasonData
goalieData <- goalieRaw %>% 
  select(4,3,losses,wins)

# skaterData is for knowing the difference in skater of specific franchise considering teamtotalData and seasonData
skaterData <- skaterRaw %>% 
  select(5,4,mostAssistsOneSeason,mostPenaltyMinutesOneSeason)

# Viewing the data after selecting variables and creating new variables
# datatable(franchiseData)
# datatable(teamtotalData)
# datatable(seasonData)
# datatable(goalieData)
# datatable(skaterData)
```

# Exploratory Data Analysis

## winPct & futureWin in teamtotalData

### Treat the winPct as categorical data

I create a new variable called winPct in teamtotalData, if winPct is equal or greater than 0.5, which means the total number of wins is greater than other kinds of situations (losses or ties), we can conclude that those franchises are more likely to win in the future. Thus, I would create a new variable called futureWin which is an indicator according to whether winPct for each franchise is less than 0.5.

$$futureWin=
\begin{cases}
1,& winPct \ge 0.5 \\
0,& \text{O.W.}
\end{cases}$$

```{r}
# create indicator named futureWin
teamtotalData2 <- teamtotalData %>%
  mutate(futureWin=ifelse(winPct<0.5,0,1))
```

According to the Contingency table of futureWin, we can know that there are three franchises greater than or equal to 0.5, which means those three franchises are more likely to win in the future. Also, those three franchises are active. And the number of active franchises is more than that of inactive franchises.

```{r}
# Contingency table for the combination of futureWin and activeFranchise
kable(table(teamtotalData2$futureWin,teamtotalData2$activeFranchise),caption="Contingency table of futureWin")

# side-by-side bar chart for the combination of futureWin and activeFranchise
ggplot(teamtotalData2,aes(x=as.factor(futureWin))) +
  geom_bar(aes(fill=as.factor(activeFranchise)),position = "dodge")+
  labs(x="futureWin",title="side-by-side bar chart for futureWin")+
  scale_fill_discrete("activeFranchise")
```

To sort out the best three franchises and the worst three franchises according to the winPct, I arrange the data sets and combine them with franchiseData, which help us to connect the franchiseId with its name. From the Worst three table of futurnWin, we know that the worst three franchises are not active, which might be the reason for their winPct. Besides, the best three franchiseId are 1,16,38, and the worst three franchiseId are 2,9,13.

```{r}
# get the best three franchises according to the winPct
fBest3 <- teamtotalData2 %>%
  arrange(desc(winPct)) %>%
  head(n=3)

# get the worst three franchises according to the winPct
fWorst3 <-teamtotalData2 %>%
  arrange(winPct) %>%
  head(n=3)

# combine above tables with franchiseData
kable(inner_join(franchiseData,fBest3),caption = "Best three table of futurnWin")
kable(inner_join(franchiseData,fWorst3),caption = "Worst three table of futurnWin")
```

### Treat the winPct as numeric data

From the summaries, we can know that the mean of winPct is 0.4326, which is less than 0.5, which means most values of winPct are less than 0.5. According to the histogram, we can see that the distribution of winPct is left-skewed. From the side-by-side boxplot, we can notice that, the winPct of seven inactives franchises are lesser than that of active franchises.

```{r}
# summaries of winPct
summary(teamtotalData$winPct)

# histogram of winPct
ggplot(teamtotalData,aes(x=winPct))+
  geom_histogram(aes(y=..density..),bins=10)+
  geom_density(size=2,color="steelblue")+
  labs(title="histogram of winPct")

# side-by-side boxplot for winPct according to activeFranchise
ggplot(teamtotalData,aes(x= as.factor(activeFranchise),y=winPct))+
  geom_boxplot()+
  geom_point(aes(color=as.factor(activeFranchise)),position = "jitter") +
  labs(x="activeFranchise", title = "side-by-side boxplot for winPct")+
  scale_fill_discrete("activeFranchise")

```

## mfWinRatio & steadyWin in seasonData

### Treat the winPct as categorical data

I also create a new variable called mfWinRatio in seasonData, if mfWinRatio more equal to 1, which means, the total number of the most wins are more closer to that of the fewest wins. However, there are some missing values, which need to be picked out considering not knowing the true story behind the those missing values. Thus, I would drop those missing records first, them I would create a new variable called steadyWin which is a indicator according to whether mfWinRatio for each franchise is closer to 1.

$$steadyWin=
\begin{cases}
1,& mfWinRatio < 2 \\
0,& \text{O.W.}
\end{cases}$$

```{r}
seasonData2 <- seasonData %>%
  # drop the missing records
  filter(!is.na(mfWinRatio)) %>%
  # create indicator named steadyWin
  mutate(steadyWin=ifelse(mfWinRatio < 2,1,0))
```

According to the Contingency table of steadyWin, we can know that there are three franchises less than 2, which means those three franchises are playing more stably. Also, those three franchises are active. And the number of active franchises is more than that of inactive franchises.

```{r}
# create Contingency table of steadyWin
kable(table(seasonData2$steadyWin),caption="Contingency table of steadyWin")

teamtotalData3<- teamtotalData2 %>%
  select(franchiseId,activeFranchise)

seasonData3 <- inner_join(seasonData2,teamtotalData3)

# side-by-side bar chart for the combination of steadyWin and activeFranchise
ggplot(seasonData3,aes(x=as.factor(steadyWin))) +
  geom_bar(aes(fill=as.factor(activeFranchise)),position = "dodge")+
  labs(x="steadyWin",title="side-by-side bar chart for steadyWin")+
  scale_fill_discrete("activeFranchise")
```

To sort out the best three franchises and the worst three franchises according to the mfWinRatio, I arrange the data sets and combine them with franchiseData, which help us to connect the franchiseId with its name. Besides, the best three franchiseId are 34,37,38, and the worst three franchiseId are 24,28,30. However, all of the most franchises are active.

```{r}
# get the worst three franchises according to the mfWinRatio
sWorst3 <- seasonData3 %>%
  arrange(desc(mfWinRatio)) %>%
  head(n=3)

# get the best three franchises according to the mfWinRatio
sBest3 <-seasonData3 %>%
  arrange(mfWinRatio) %>%
  head(n=3)

# combine above tables with franchiseData
kable(inner_join(franchiseData,sBest3),caption = "Best three table of steadyWin")
kable(inner_join(franchiseData,sWorst3),caption = "Worst three table of steadyWin")
```


### Treat the winPct as numeric data

From the summaries, we can know that the mean of mfWinRatio is 3.327, which is greater than 1, which means most values of mfWinRatio are greater than 1. According to the histogram, we can see that the distribution of mfWinRatio is right-skewed. From the side-by-side boxplot, we can notice that, except those missing records, only one inactive franchise exists in the seasonData, which might be the season why this record seems more steady than other data points.

```{r}
# summaries of mfWinRatio
summary(seasonData$mfWinRatio)

# histogram of mfWinRatio
ggplot(seasonData,aes(x=mfWinRatio))+
  geom_histogram(aes(y=..density..),bins=10)+
  geom_density(size=2,color="steelblue")+
  labs(title="histogram of mfWinRatio")

# side-by-side boxplot for mfWinRatio according to activeFranchise
ggplot(seasonData3,aes(x= as.factor(activeFranchise),y=mfWinRatio))+
  geom_boxplot()+
  geom_point(aes(color=as.factor(activeFranchise)),position = "jitter") +
  labs(x="activeFranchise", title = "side-by-side boxplot for mfWinRatio")+
  scale_fill_discrete("activeFranchise")
```

## The best and Worst id
```{r}
inner_join(fBest3,sBest3) %>%
  select(franchiseId,winPct,mfWinRatio)
newfWorst3 <- teamtotalData2 %>%
  filter(activeFranchise==1)%>%
  arrange(winPct) %>%
  head(n=3)
inner_join(newfWorst3,sWorst3) %>%
  select(franchiseId,winPct,mfWinRatio)
```
## The difference in goalie of 38 and 28
```{r}
goalieData2 <- goalieData %>%
  filter(franchiseId==38 |franchiseId==28)
skaterData2 <- skaterData %>%
  filter(franchiseId==38 |franchiseId==28)
```

```{r}
ggplot(goalieData2,aes(x=losses ,y=wins,group=as.factor(franchiseId)))+
  geom_point(aes(color=as.factor(franchiseId)))+
  geom_smooth(method = lm, colour="green")+
  labs(title = "losses vs wins")
ggplot(skaterData2,aes(x=mostAssistsOneSeason ,y=mostPenaltyMinutesOneSeason,group=as.factor(franchiseId)))+
  geom_point(aes(color=as.factor(franchiseId)))+
  geom_smooth(method = lm, colour="blue")+
  labs(title = "goals vs points")
```


# Conclusion

# Discussion on this project
## what would you do differently?
## what was the most difficult part for you?
## what are your big take-aways from this project?
